---
title: "week_7_hw"
author: "Sofya Lipkind"
date: "6/2/2020"
output: 
  html_document: 
    highlight: pygments
    keep_md: yes
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Chapter 13 Problems  

### 13.2.1

#### 1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

*airport contains the variables Lat and Lon, which describe the location of each airport.Using Lat an Lon in conjunction with origin and dest (from flight) should give us a vague idea of the path between the origin and the destination.*

#### 2. I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

*weather(origin) -> airports(faa) -> airports(names)*

#### 3. weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

*dest, correct?*

#### 4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

```{r}
flights %>%
  group_by(year) %>%
  count(month, day) %>%
  arrange(n)

#connect via year/month/day to flights, whereby it could connect to other tables
```

### 13.3.1

#### 1. Add a surrogate key to flights.
```{r}
flights %>%
  mutate(id=row_number())
```
#### 2. Identify the keys in the following datasets
  1. Lahman::Batting  
```{r}
head(Lahman::Batting)
#primary: playerID
```

  2. babynames::babynames  
```{r}
glimpse(babynames::babynames)
unique(babynames::babynames$year)
#primary: name + year
```

### 13.4.6

#### 1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
#could not install maps
#install.packages("maps")
#airports %>%
#  semi_join(flights, c("faa" = "dest")) %>%
#  ggplot(aes(lon, lat)) +
#    borders("state") +
#    geom_point() +
#    coord_quickmap()
```

#### 2. Add the location of the origin and destination (i.e. the lat and lon) to flights.

```{r}

```

## Chapter 18 Problems

