---
title: "Week 6 HW"
author: "Sofya Lipkind"
date: "5/20/2020"
output: 
  html_document: 
    highlight: pygments
    keep_md: yes
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r problem tables, eval=FALSE, include=FALSE}
table1
#> # A tibble: 6 x 4
#>   country      year  cases population
#>   <chr>       <int>  <int>      <int>
#> 1 Afghanistan  1999    745   19987071
#> 2 Afghanistan  2000   2666   20595360
#> 3 Brazil       1999  37737  172006362
#> 4 Brazil       2000  80488  174504898
#> 5 China        1999 212258 1272915272
#> 6 China        2000 213766 1280428583
table2
#> # A tibble: 12 x 4
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362
#> # … with 6 more rows
table3
#> # A tibble: 6 x 3
#>   country      year rate             
#> * <chr>       <int> <chr>            
#> 1 Afghanistan  1999 745/19987071     
#> 2 Afghanistan  2000 2666/20595360    
#> 3 Brazil       1999 37737/172006362  
#> 4 Brazil       2000 80488/174504898  
#> 5 China        1999 212258/1272915272
#> 6 China        2000 213766/1280428583

# Spread across two tibbles
table4a  # cases
#> # A tibble: 3 x 3
#>   country     `1999` `2000`
#> * <chr>        <int>  <int>
#> 1 Afghanistan    745   2666
#> 2 Brazil       37737  80488
#> 3 China       212258 213766
table4b  # population
#> # A tibble: 3 x 3
#>   country         `1999`     `2000`
#> * <chr>            <int>      <int>
#> 1 Afghanistan   19987071   20595360
#> 2 Brazil       172006362  174504898
#> 3 China       1272915272 1280428583
```

## Chapter 12 Problems

### 12.2.1

 2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations: 
 
    + Extract the number of TB cases per country per year.  
    + Extract the matching population per country per year.  
    + Divide cases by population, and multiply by 10000.  
    + Store back in the appropriate place.  
```{r}
#table2
data(table2)
table2_cases <- table2 %>%
  spread(type, count) %>%
  mutate(rate = cases / population * 10000)
print(table2_cases)
```
```{r}
#table4a + 4b = 4c

table4 <- inner_join(table4a, table4b, by = "country") %>%
  gather(year, cases, c(`1999.x`, `2000.x`)) %>%
  gather(year, population, c(`1999.y`, `2000.y`)) %>%
  mutate(rate = cases / population * 10000) %>%
#  gsub(".y","", table4$year) %>%      Why does including this line get rid of all other information?
  print()

#gsub(".y","", table4$year)
```

Which representation is easiest to work with? Which is hardest? Why?  
*Both of them sucked. How are you supposed to do these without knowing gather() and spread()?*
*I suppose Table 4 was harder, since I had to join them before I could do anything else, and then separate . Then again, if I didn't have spread, I would have hated 2 a lot, as well.*

 3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?
```{r}
#> # A tibble: 12 x 4
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362
#> # … with 6 more rows

#option 1: without using tidied data frame
table2.3 <- table2 %>%
  filter(type == "cases") #need to filter out population first

ggplot(table2.3, aes(year, count)) +
  geom_line(aes(group = country)) +
  geom_point(color  = country)
```
```{r}
#option 2: using tidied data frame
table2_cases %>% ggplot(aes(year, cases)) +
  geom_line(aes(group = country)) +
  geom_point(aes(color  = country))
```

### 12.3.3

 1. Why are pivot_longer() and pivot_wider() not perfectly symmetrical? Carefully consider the following example:
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
print(stocks)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
```
*Oh, that's interesting! In the original tibble (stocks), the `year` column is of class **double**, whereas once the OG tibble has been passed through wider -> longer, `year` is now of class **character**.*
*If one wanted to return `year` to the prior state, you would have to perform the extra step of parsing it.*
pivot_longer() has a names_ptype argument, e.g. names_ptype = list(year = double()). What does it do?
```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptypes = list(year = double()))
```
*It apparently eliminates the need to parse after performing a round of ~~spreading~~ pivoting wider and ~~gathering~~ pivoting longer.*

 2. Why does this code fail?  
```{r}
#table4a %>% 
#  pivot_longer(c(1999, 2000), names_to = "year", values_to = "cases")
```
*You forgot the tickmarks for non-syntactic column names, ya dingus!*
```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

 3. What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?
```{r}
people <- as_tibble(tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45, #age 1
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50, #age 2
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
))
# oh dang, a tribble, f that noise, I don't want to deal with formulas
# Phillip Woods has two ages??

people %>%
  group_by(name,names) %>%
  mutate(count = count(names)) %>%
  print()

people_height <- as_tibble(people) %>%
  unite(names_values,c(names,values), sep = "_", na.rm = FALSE) %>%
  spread(name,names_values) %>%
 # pivot_longer(c(age_37,age_45,age_50), names_to = "age", values_to = "age_name") %>%
  #pivot_longer(c(height_156,height_186), names_to = "height", values_to = "height_name") %>%
  print()
  
```

 4. 

### 12.4.3

 1. 

 2. 

 3. 

### 12.6.1

 3. 

 4. 

