---
title: "Week 6 HW"
author: "Sofya Lipkind"
date: "5/20/2020"
output: 
  html_document: 
    highlight: pygments
    keep_md: yes
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r problem tables, eval=FALSE, include=FALSE}
table1
#> # A tibble: 6 x 4
#>   country      year  cases population
#>   <chr>       <int>  <int>      <int>
#> 1 Afghanistan  1999    745   19987071
#> 2 Afghanistan  2000   2666   20595360
#> 3 Brazil       1999  37737  172006362
#> 4 Brazil       2000  80488  174504898
#> 5 China        1999 212258 1272915272
#> 6 China        2000 213766 1280428583
table2
#> # A tibble: 12 x 4
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362
#> # … with 6 more rows
table3
#> # A tibble: 6 x 3
#>   country      year rate             
#> * <chr>       <int> <chr>            
#> 1 Afghanistan  1999 745/19987071     
#> 2 Afghanistan  2000 2666/20595360    
#> 3 Brazil       1999 37737/172006362  
#> 4 Brazil       2000 80488/174504898  
#> 5 China        1999 212258/1272915272
#> 6 China        2000 213766/1280428583

# Spread across two tibbles
table4a  # cases
#> # A tibble: 3 x 3
#>   country     `1999` `2000`
#> * <chr>        <int>  <int>
#> 1 Afghanistan    745   2666
#> 2 Brazil       37737  80488
#> 3 China       212258 213766
table4b  # population
#> # A tibble: 3 x 3
#>   country         `1999`     `2000`
#> * <chr>            <int>      <int>
#> 1 Afghanistan   19987071   20595360
#> 2 Brazil       172006362  174504898
#> 3 China       1272915272 1280428583
```

## Chapter 12 Problems

### 12.2.1

 2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations: 
 
    + Extract the number of TB cases per country per year.  
    + Extract the matching population per country per year.  
    + Divide cases by population, and multiply by 10000.  
    + Store back in the appropriate place.  
```{r}
#table2
data(table2)
table2_cases <- table2 %>%
  spread(type, count) %>%
  mutate(rate = cases / population * 10000)
print(table2_cases)
```
```{r}
#table4a + 4b = 4c

table4 <- inner_join(table4a, table4b, by = "country") %>%
  gather(year, cases, c(`1999.x`, `2000.x`)) %>%
  gather(year, population, c(`1999.y`, `2000.y`)) %>%
  mutate(rate = cases / population * 10000) %>%
#  gsub(".y","", table4$year) %>%      Why does including this line get rid of all other information?
  print()

#gsub(".y","", table4$year)
```

Which representation is easiest to work with? Which is hardest? Why?  
*Both of them sucked. How are you supposed to do these without knowing gather() and spread()?*
*I suppose Table 4 was harder, since I had to join them before I could do anything else, and then separate . Then again, if I didn't have spread, I would have hated 2 a lot, as well.*

 3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?
```{r}
#> # A tibble: 12 x 4
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362
#> # … with 6 more rows

#option 1: without using tidied data frame
table2.3 <- table2 %>%
  filter(type == "cases") #need to filter out population first

ggplot(table2.3, aes(year, count)) +
  geom_line(aes(group = country)) +
  geom_point(aes(color  = country))
```
```{r}
#option 2: using tidied data frame
table2_cases %>% ggplot(aes(year, cases)) +
  geom_line(aes(group = country)) +
  geom_point(aes(color  = country))
```

### 12.3.3

 1. Why are pivot_longer() and pivot_wider() not perfectly symmetrical? Carefully consider the following example:
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
print(stocks)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
```
*Oh, that's interesting! In the original tibble (stocks), the `year` column is of class **double**, whereas once the OG tibble has been passed through wider -> longer, `year` is now of class **character**.*
*If one wanted to return `year` to the prior state, you would have to perform the extra step of parsing it.*  
pivot_longer() has a names_ptype argument, e.g. names_ptype = list(year = double()). What does it do?
```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptypes = list(year = double()))
```
*It apparently eliminates the need to parse after performing a round of ~~spreading~~ pivoting wider and ~~gathering~~ pivoting longer.*

 2. Why does this code fail?  
```{r}
#table4a %>% 
#  pivot_longer(c(1999, 2000), names_to = "year", values_to = "cases")
```
*You forgot the tickmarks for non-syntactic column names, ya dingus!*
```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

 3. What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value? *(absolute nightmare question)*
```{r}
people <- as_tibble(tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45, #age 1
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50, #age 2
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
))

f_it <- as.data.frame(people)
f_it[3,2] = "age2"
print(f_it)

f_it <- as_tibble(f_it) %>%
  spread(names,name) %>%
  print()

#people %>%
#  group_by(names) %>%
#  mutate(group_id <- group_indices(values)) %>%
#  print()
```

 4. Tidy the simple tibble below. Do you need to make it wider or longer? What are the variables?
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

preggo <- as_tibble(preg) %>%
  pivot_longer(c("male","female"), names_to = "gender", values_to = "count") %>%
  select(gender, pregnant, count) %>%
  print()
```
 

### 12.4.3

 1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.
```{r}
#extra options : warn(default), drop, merge(length(into))
##fill options : warn(emit warning and fill from right), right(fill with missing values on right), left(fill with missing vlues on left)

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% #one row has an extra data point
  separate(x, c("one", "two", "three"))
#becomes
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% #one row has an extra data point
  separate(x, c("one", "two", "three"), extra = "merge")
#fill makes no difference here b/c no empty values
```
```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%  #one row lacks a data point
  separate(x, c("one", "two", "three"))
#becomes
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%  #one row lacks a data point
  separate(x, c("one", "two", "three"), fill = "left") #NA shifted to right (no effect if fill = "right)
```

 2. Both unite() and separate() have a remove argument*(defaults to TRUE)*. What does it do? Why would you set it to FALSE?
```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%  #one row lacks a data point
  separate(x, c("one", "two", "three"), remove = FALSE)
```
*`remove` allows one to remove a column describing the contents of each row of the original tibble.*  
*Ordinarily, this is useful to remove unnecessary clutter.*  
*Setting `remove = FALSE` could be helpful for someone to check if they separated their observations correctly.*  

 3. Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?  
 *`extract()` appears to separate with an additional option to separate according to a regex pattern, which could allow one to cleave cells with a complex pattern that couldn't be caught with `separate()`.*
 *If one had a sentence as an observation, they could separate it in any number of ways (by each letter, each space, by clauses, by SOV), but one can only unite the sentence correctly one way.*  
### 12.6.1
```{r}
#relevant dataset
data(who)
whom <- who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) %>%
  print()
```


 3. I claimed that iso2 and iso3 were redundant with country. Confirm this claim.
```{r}
#count unique values for each column (status = weak evidence)
  summary(unique(who$country)) #219
  summary(unique(who$iso2)) #219
  summary(unique(who$iso3)) #219
  
#unique combinations of country with iso2 and iso3
#would expect to find 219 combinations if each country has the same equivalent in iso2 and iso3 (status = slightly better evidence!)

nrow(unique(who[,c("country","iso2","iso3")])) #219
```
 
 4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.
 
 country > year > sex
```{r}
#did not get to in time
# whom %>%
# group_by(country,year,sex) %>%
# count(cases)
```
 

