---
title: "Week 2 HW"
author: "Sofya Lipkind"
date: "4/22/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)
```

##CHAPTER 3: VISUALIZATIONS

```{r}
picsave <- function(graph, name) {
  ggsave(plot = graph, filename= name, device = "pdf", width = 12, height = 8, path = "~/GitHub/S_Lipkind_Rundergrad2020/week1/")
}
```


###Q: Do cars w/ big engines use more fuel than w/ small engines? (dataframe *mpg*)
_note that this was done without the textbook, so it may not follow those results strictly. Just wanted to see how I would approach answering the question._
```{r}
head(mpg)
summary(mpg)
#manufacturer model ,displ(engine size in L),  year,   cyl(#cylinders), trans(manual vs auto),  drv(front, rear, or 4wd),     cty(city mpg),   hwy(fuel efficiency in mpg), fl(fuel type),    class(category of car)  
#What variables seem relevant to this question?
# displ, hwy, possibly cty


plot( mpg$hwy ~ mpg$displ)
#inverse relationship, roughly - as engine size increases, fuel efficiency decreases
plot(mpg$cty ~ mpg$displ)
#slightly steeper inverse relationship


#well... that's cool, but let's do it in ggplot. Can I put both cty and hwy on the same plot?

plots1 <- mpg %>% ggplot(
  aes(x = displ)) +
  geom_jitter(aes(y = cty), color = "maroon", shape = 8) +
  geom_jitter(aes(y = hwy), color = "cornflowerblue", shape = 1) + #REVELATION TO JUST DO TWO GEOM_JITTERS!!! Bless you, Stack Overflow.
  ggtitle("Comparing mpg in cities vs highways by engine size") + 
  scale_x_continuous(name = "engine size (L)") +
  scale_y_continuous(name = "fuel efficiency (mpg)")
plots1
plots1 + annotate(label = "Makeshift legend: maroon = city, blue = highway", x = 5, y = 50, geom = "text") #QUESTION: How would I get the legend to work when I manually set the colors and shapes b/w two geom_jitters?

#Hooray! I mean, it doesn't help me quantify things any more, but it makes it easier to compare cty vs hwy

picsave(plots1, "warmupplot.pdf")

#How might we quantify this relationship a bit more?

city <- lm(data = mpg, cty ~ displ)
summary(city)

highway <- lm(data = mpg, hwy ~ displ)
summary(highway)

#conclusions:
# 1. highway had 10 more mpg than city, without engine size effect
# 2. highway was slightly more negatively affected by engine size (-3.5 vs -2.6)
# 3. these results are highly significant
```


### QUESTION 3.2.4 (1-5)

1: Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```
I see a blank image.

2: How many rows are in mpg? How many columns?

There are 234 observations (rows), and 11 columns. 

3:What does the drv variable describe? Read the help for ?mpg to find out.

drv = is the vehicle:  
* front-wheel drive  
* back-wheel drive  
* four-wheel drive

4:Make a scatterplot of hwy vs cyl.

```{r}
plot(data = mpg, cyl ~ hwy)

#makes sense, since cyl is a discrete numerical variable, and hwy is continuous numerical
```

5:What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
#plot(data = mpg, drv ~ class)
```

Well, class and drv are both categorical variables, so they are unsuited to the quantitative lens given by a scatterplot. Since there is no quantitative data for plot() to work with, there is an ERROR.


###QUESTION 3.3.1 (1-6)

1:What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Well, you dingus, you put your closing aes parenthesis in the wrong spot! Color is assigned as an aesthetic with a dummy variable ("blue"), so it just does whatever.

If you were to instead do this:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

See? Much better!

2: Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
labels(mpg)
#shows column titles
summary(mpg)
#shows all columns, along with class and any other relevant information
#Heuristic:
#Categorical variables will show as a 'character' class (or as factors, iirc).
#Continuous variables will show a quantile distribution.
#
#Exceptions: dates or times are categorical, such as 'year'; variables that can only have whole numbers (e.g. cyl) are categorical.
```

3:Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
mpg
contvariable <- mpg %>% ggplot(
  aes(x = cty)) +  #ERROR: not finding cty for some reason?
    geom_bar(color = cty, size = cty, shape = cty) #not enough shape options for a continuous variable, possibly, thereby causing the error?

contvariable
catvariable <- mpg %>% ggplot(
  aes(x = cyl)) +
    geom_bar(color = cyl, size = cyl, shape = cyl)
catvariable

```

4: What happens if you map the same variable to multiple aesthetics?

The variable will be distinguished by multiple aesthetics. See: plot1

5: What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
mpg %>% ggplot(
  aes(y = hwy,
      x = cyl)) +
  geom_point(shape = 21, stroke = .1)

#stroke determines border thickness of shapes that have separate fills and borders

```

6: What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
mpg %>% ggplot(
  aes(x = hwy,
      y = cyl,
      colour = displ < 5)) +
  geom_point()

#if/else statement. If variable meets condition, TRUE, if not, FALSE. The two will be delineated
```


###QUESTION 3.5.1 (1-6)

1. What happens if you facet on a continuous variable?

```{r}
mpg %>% ggplot(
  aes(x = drv, y = displ)) +
    geom_point() +
  facet_grid(displ ~ .)
#answer: chaos ensues
```


2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
#blank space == there is no row that contains that specific combination of drv and cyl
```

3. What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

#. in the context of facetting, excludes that position (x,y) from being facetted; i.e. they'll be grouped in that dimension
```


4.Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

  If there is significant overlap to the points, or otherwise general busyness(perhaps due to large dataset), facetting can help make things clearer by breaking up the data. However, it makes it harder to visually compare the facetted qualities

5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

nrow = number of rows for facets
ncol = number of columns for facets
together = shape of grid to be formed

facet_grid is used for more complex data where multiple dimensions are being facetted, not just one. Therefore, the number of rows and columns is significant.

6.When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

More legible?


###QUESTION 3.6.1 (1-5, one graph from 6)

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_line()
geom_boxplot()
geom_histogram()
geom_area()


2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
What does show.legend = FALSE do? What happens if you remove it?  

show.legend = FALSE removes the legend for drv categories that would otherwise appear on the right. 

Why do you think I used it earlier in the chapter?  

It was unnecessary at that point.

3. What does the se argument to geom_smooth() do?

Shows standard error for each trendline.

4. Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
  
  They look the same. The first plot just contains the variables in the beginning section, so they don't need to be defined for subsequent geoms. 
  
  
5. Recreate the R code necessary to generate the following graphs.