---
title: "Week 4 HW"
output:
  html_document: 
    keep_md: yes
---

## SETUP
```{r setup, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
library(tidyverse)
setwd("~/GitHub/S_Lipkind_Rundergrad2020/week4/")
```

FXN: Save graphical outputs
```{r}
picsave <- function(graph, name) {
  ggsave(plot = graph, filename= name, device = "pdf", width = 12, height = 8, path = "~/GitHub/S_Lipkind_Rundergrad2020/week4/pics/")
}
```
---
## CHAPTER 7 PROBLEMS

### 7.3.4

#### 1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
glimpse(diamonds)
#(xyz) are continuous variables, measuring the dimensions of each diamond
#Want to see distributions of each axis, but hmm, the three axes are not united by anything that I can easily call in ggplot in one go
xyzdist <- ggplot(diamonds, mapping = aes()) +
  geom_histogram(x, position = "dodge", color = IQR(x)) +
  geom_histogram(y, position = "dodge", color = IQR(y)) +
  geom_histogram(z, position = "dodge", color = IQR(z))
xyzdist
```


#### 2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

#### 3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

#### 4. Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?




### 7.4.1

#### 1.

#### 2.


### 7.5.1.1

#### 1.

#### 2.

#### 3.

#### 5.

#### 6.


### Last Problem

#### "Describe an *experiment* that would possess 2 *categorical* variables and 2 *continuous* variables."

