---
title: "Week 3 HW and Notes"
output:
  html_document: 
    keep_md: yes
editor_options:
  chunk_output_type: console
---
# SETUP
```{r setup, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
library(tidyverse)
library(nycflights13)
setwd("~/GitHub/S_Lipkind_Rundergrad2020/week3/")
```

FXN: Save graphical outputs
```{r}
picsave <- function(graph, name) {
  ggsave(plot = graph, filename= name, device = "pdf", width = 12, height = 8, path = "~/GitHub/S_Lipkind_Rundergrad2020/week3/pics/")
}
```
---
# PROBLEMS

## CHAPTER 4 PROBLEMS

### 4.4

1. Why does this code not work?

```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

*Was this written by someone who speaks Turkish or something? Not sure how else someone could use _ı_ instead of _i_.*

2. Tweak each of the following R commands so that they run correctly:

```{r}
a <- ggplot(data = mpg) + #dota -> data
  geom_point(mapping = aes(x = displ, y = hwy))
#picsave(a, "4.4.2 graph.pdf")

filter(mpg, cyl == 8) #= -> ==
filter(diamonds, carat > 3) # diamond -> diamonds

```

3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

*Oh my goodness, that is amazing. An entire list of keyboard shortcuts. You could also reach that page by going to Help > Keyboard Shortcuts Help.*

## CHAPTER 5 PROBLEMS

### 5.2.4

Find all flights that...

  #### 1.1: had an arrival delay of two or more hours
  
```{r}
head(flights)
filter(flights, arr_delay >= 2)
```
  
  #### 1.4: Departed in summer (July, August, and September)
  
```{r}
filter(flights, month %in% c(7,8,9))
```

  #### 1.5: Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights, arr_delay > 2 & dep_delay == 0)
```

  #### 1.7: Departed between midnight and 6am (inclusive)
  
```{r}
(d <- filter(flights, dep_time <= 2400, dep_time >= 600))
```
  
  #### 2: Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

?between
*It's an inclusive shortcut to find values within a certain range.*
```{r}
e <- filter(flights, dep_time %in% between(dep_time,0, 600))
# d == e
```

  #### 3: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, dep_time %in% NA) # 8255 rows/flights

#these observations also all have NA dep_delay, arr_time, arr_delay.
#Hypothesis: cancelled flights
```

  
### 5.3.1

#### 1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_delay))) #this works, but is it the intended solution?
```

#### 2.Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, dep_time, desc(dep_delay))
```

#### 3. Sort flights to find the fastest (highest speed) flights.
```{r}
y <- arrange(flights, desc(distance), air_time)
#(select(y, distance, air_time)) -> double-checking
```

#### 4. Which flights travelled the farthest? Which travelled the shortest?
```{r}
(longest_distance <- top_n(flights, 10, distance))
(shortest_distance <- top_n(flights, -10, distance))
```
 
 
### 5.4.1

#### 1.Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
#one option: select()
```

#### 2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, day, month, day, month, dep_delay, dep_delay)
```
*Repeating variable names does not appear to make a difference. Only the sorting of the initial appearance of each name within the list matters.*

#### 3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay") 
select(flights, one_of(vars))
```

*one_of() allows one to make a character vector with specific column names that you can then select for.*

#### 4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))
```

*The results aren't too surprising, though I didn't realize select was not case-sensitive.*
*If I wanted to specify case, I could add the specifier below:*
```{r}
select(flights, contains("time", ignore.case = FALSE))
```


### 5.5.2

#### 1.Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
head(flights)

minutes <- function(miltime) {
  miltime %/% 100 * 60 + miltime %% 100
}

mod <- flights %>%
  select(dep_time,sched_dep_time) %>%
  lapply(.,minutes) %>%
  as_tibble() %>%
  rename(
    mod_dep_time = dep_time,
    mod_sched_dep_time = sched_dep_time)

modflights <- flights %>% #I initially tried to join flights and mod together, but I couldn't figure out the 'by' part.
  mutate(
    mod_dep_time = mod$mod_dep_time,
    mod_sched_dep_time = mod$mod_sched_dep_time)
```

#### 2.Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r echo=TRUE}
discrepancy <- flights %>%
  mutate(
    arr_dep = arr_time - dep_time) %>%
  select(arr_dep, air_time) %>%
  lapply(.,minutes) %>%
  glimpse(.)

#Theoretically, you would expect the values to be the same, but they are not.
#Perhaps there's an issue as a result of military time? When I converted to minutes, though, the difference remained. Air time is shorter than arr_dep.
#Hypothesis: The difference in time is explained by the time the planes take to embark and disembark.
#How to test: add dep_delay and arr_delay to ait_time, see if air_time == arr_dep.
#To self: Dingus, you needed o convert military time to minutes ahead of time, before mutating

head(flights)

flights %>% 
    mutate(
      air_time_delay = air_time + arr_delay + dep_delay,
      arr_dep = discrepancy$arr_dep) %>%
    select(air_time,air_time_delay,arr_dep, dep_delay, arr_delay) %>%
    lapply(., minutes) %>%
    glimpse(.)

#However, that doesn't seem to work, either, and the difference between arr_dep and air_time is not explained neatly by either dep_delay or arr_delay.
```

#### 3.Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
```{r}
#Expected: dep_time == sched_dep_time + dep_delay
#
flights %>%
  select(dep_time, sched_dep_time, dep_delay) %>%
  mutate(hypothesis = sched_dep_time + dep_delay) %>%
  lapply(., minutes) %>%
  glimpse(.)

#Hypothesis is correct!
```

#### 4.Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().
```{r}
#solution 1: top_n()

flights %>%
  mutate(
    delays4dayz = dep_delay + arr_delay) %>%
  top_n(.,10, delays4dayz) %>%
  glimpse(.)

#If one were to instead use min_rank(), ties would be settled by choosing the minimum of the "corresponding indices," whatever that means.
```

#### 5.What does 1:3 + 1:10 return? Why?
```{r}
1:3 + 1:10
#Returns:   2  4  6  5  7  9  8 10 12 11
```
*Because 1:3 is shorter than 1:10, 1:3 "loops" as it adds along 1:10.*  
*(1 2 3 1 2 3 1 2 3 1) + (1 2 3 4 5 6 7 8 9 10)*

#### 6.What trigonometric functions does R provide?
```{r}
sin(pi)
cos(pi)
tan(pi)
#sec(pi) Not included
#csc(pi) Not included
#cot(pi) Not included
```

### 5.6.7

#### 5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))
```{r}
#Carrier that has longest delays, on average
glimpse(flights)
flights %>%
  group_by(carrier, dest) %>%
  summarize(
    ad = mean(arr_delay, na.rm = TRUE), 
    dd = mean(dep_delay, na.rm = TRUE),
    n()) %>%
  arrange(desc(ad + dd))
#   carrier dest     ad    dd `n()`
#   <chr>   <chr> <dbl> <dbl> <int>
# 1 UA      STL   110    77.5     2
# 2 OO      ORD   107    67       1
# 3 OO      DTW    68.5  61       2
# 4 UA      RDU    56    60       1
# 5 EV      PBI    40.7  48.7     6
# 6 EV      TYS    41.2  41.8   323
# 7 EV      CAE    42.8  36.7   113
# 8 EV      TUL    33.7  34.9   315
# 9 EV      OKC    30.6  30.6   346
#10 UA      JAC    29.9  28.7    23
```
```{r}
#Now, how would I handle distinguishing bad airports vs bad carriers? 
#Bad airports will affect any airplanes that (dis)embark there, and bad airlines will affect all of their associated airplanes.
flights %>%
 group_by(carrier, dest) %>%
 summarize(
   ad = mean(arr_delay, na.rm = TRUE),
   dd = mean(dep_delay, na.rm = TRUE),
   count = n()) %>%
 arrange(desc(count))
```

