---
title: "Week 3 HW and Notes"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---
# SETUP
```{r include=FALSE}
library(tidyverse)
library(nycflights13)
setwd("~/GitHub/S_Lipkind_Rundergrad2020/week3/")
```

FXN: Save graphical outputs
```{r}
picsave <- function(graph, name) {
  ggsave(plot = graph, filename= name, device = "pdf", width = 12, height = 8, path = "~/GitHub/S_Lipkind_Rundergrad2020/week3/pics/")
}
```

# NOTES

## CHAPTER 4
  N/A
  
## CHAPTER 5

```{r include=FALSE}
flights <- flights #so I can see it in my Environment
```

  
# PROBLEMS
## CHAPTER 4 PROBLEMS

### 4.4

1. Why does this code not work?

```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

_Was this written by someone who speaks Turkish or something? Not sure how else someone could use *ı* instead of *i*._

2. Tweak each of the following R commands so that they run correctly:

```{r}
a <- ggplot(data = mpg) + #dota -> data
  geom_point(mapping = aes(x = displ, y = hwy))
#picsave(a, "4.4.2 graph.pdf")

filter(mpg, cyl == 8) #= -> ==
filter(diamonds, carat > 3) # diamond -> diamonds

```

3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

*Oh my goodness, that is amazing. An entire list of keyboard shortcuts. You could also reach that page by going to Help > Keyboard Shortcuts Help.*

## CHAPTER 5 PROBLEMS

### 5.2.4

Find all flights that...

  #### 1.1: had an arrival delay of two or more hours
  
```{r}
head(flights)
filter(flights, arr_delay >= 2)
```
  
  #### 1.4: Departed in summer (July, August, and September)
  
```{r}
filter(flights, month %in% c(7,8,9))
```

  #### 1.5: Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights, arr_delay > 2 & dep_delay == 0)
```

  #### 1.7: Departed between midnight and 6am (inclusive)
  
```{r}
(d <- filter(flights, dep_time >= 0, dep_time <= 600))
```
  
  #### 2: Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

?between
*It's an inclusive shortcut to find values within a certain range.*
```{r}
e <- filter(flights, dep_time %in% between(dep_time,0, 600))
# d == e
```

  #### 3: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, dep_time %in% NA) # 8255 rows/flights

#these observations also all have NA dep_delay, arr_time, arr_delay.
#Hypothesis: cancelled flights
```

  
### 5.3.1

#### How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_delay))) #this works, but is it the intended solution?
```

#### Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, dep_time, desc(dep_delay))
```

#### Sort flights to find the fastest (highest speed) flights.
```{r}
y <- arrange(flights, desc(distance), air_time)
#(select(y, distance, air_time)) -> double-checking
```

#### Which flights travelled the farthest? Which travelled the shortest?
```{r}
(longest_distance <- top_n(flights, 10, distance))
(shortest_distance <- top_n(flights, -10, distance))
```
 
 
### 5.4.1

#### Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
#one option: select()
```

#### What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, day, month, day, month, dep_delay, dep_delay)
```
*Repeating variable names does not appear to make a difference. Only the sorting of the initial appearance of each name within the list matters.*

#### What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay") 
select(flights, one_of(vars))
```

*one_of() allows one to make a character vector with specific column names that you can then select for.*

#### Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))
```

*The results aren't too surprising, though I didn't realize select was not case-sensitive.*
*If I wanted to specify case, I could add the specifier below:*
```{r}
select(flights, contains("time", ignore.case = FALSE))
```


### 5.5.2

#### Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
head(flights)

minutes <- function(miltime) {
  miltime %/% 100 * 60 + miltime %% 100
}

mod <- flights %>%
  select(dep_time,sched_dep_time) %>%
  lapply(.,minutes) %>%
  as_tibble() %>%
  rename(
    mod_dep_time = dep_time,
    mod_sched_dep_time = sched_dep_time) %>%
  left_join(.,flights)
mod
  
```

#### Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

#### Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

#### Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

#### What does 1:3 + 1:10 return? Why?

#### What trigonometric functions does R provide?