---
title: "Week 3 HW and Notes"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# SETUP
```{r}
library(tidyverse)
library(nycflights13)
setwd("~/GitHub/S_Lipkind_Rundergrad2020/week3/")
```

FXN: Save graphical outputs
```{r}
picsave <- function(graph, name) {
  ggsave(plot = graph, filename= name, device = "pdf", width = 12, height = 8, path = "~/GitHub/S_Lipkind_Rundergrad2020/week3/pics/")
}
```

# NOTES

##CHAPTER 4
  N/A
  
##CHAPTER 5

```{r}
flights <- flights #so I can see it in my Environment
```

  
# PROBLEMS
## CHAPTER 4 PROBLEMS

### 4.4

1. Why does this code not work?

```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

_Was this written by someone who speaks Turkish or something? Not sure how else someone could use *ı* instead of *i*._

2. Tweak each of the following R commands so that they run correctly:

```{r}
a <- ggplot(data = mpg) + #dota -> data
  geom_point(mapping = aes(x = displ, y = hwy))
picsave(a, "4.4.2 graph.pdf")

filter(mpg, cyl == 8) #= -> ==
filter(diamonds, carat > 3) # diamond -> diamonds

```

3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

*Oh my goodness, that is amazing. An entire list of keyboard shortcuts. You could also reach that page by going to Help > Keyboard Shortcuts Help.*

## CHAPTER 5 PROBLEMS

### 5.2.4

  ####1.1
  
  

  ####1.4

  ####1.5

  ####1.7

  ####2
  
  ####3
  
### 5.3.1
